Notes of Group Theory and Physics

### 2017-02-15 Lecture note - Group Theory in Physics/Group theory and Physics Sternberg/sternberg.tex
[major] Note of Sternberg's Group Theory and Physics: about the 32 point groups.

\section{The Classification of the finite subgroups of
\texorpdfstring{$\mathrm{SO}(3)$}{} (1.8)}
\label{sec:Classification-of-finite-subgroups-of-SO3}

@@

This equation can be simplified by considering the practical numerical values of
$n$, $r$ and $n_i$, with $n_i \leq n$. By eliminating case by case, he finally arrived
at five sets of possible values:
\begin{table}[H]
\centering
\label{tab:label}
\caption{Finite rotation groups}
\begin{tabular}{l l l l l l}
    \hline
    $r$ & $(n_1,n_2,n_3)$ & $\# G$ & Schoenflies & Hermann-Mauguin & Note
    \\
    \hline
    \hline
    2 & (n,n,0) & n & $C_n$ & $n$ & Cyclic group \\
    3 & $(2,2,k), k\geq 2$ & $2k$ & $D_k$ & $222$ for $D_2$, $k2$ otherwise & Dihedral group \\
    3 & $(2,3,3)$ & 12  & $T$  & 23& of regular tetrahedron \\
    3 & $(2,3,4)$ & 24 & $O$ & 432 & of regular octahedron \\
    3 & $(2,3,5)$ & 60 & $I$ & not mentioned & of icosahedron \\
    \hline
\end{tabular}
\end{table}
For details about derivation please visit pp. 28 to 31 of \cite{Sternberg1994}.

But we need to exclude some subgroups from the list of crystallographic groups.
The result is that, on the first row, $n$ is restricted to be $1,2,3,4$ and $6$;
on the second row, $k$ is restricted to be $2,3,4$ and $6$. And $I$ is excluded
from the list of crystallographic groups.

@@

\bibliography{cite}
\bibliographystyle{alpha}

>>>>

\section{Basic definitions and examples (Ch 1)}
\label{sec:Basic-def-examples}

    \subsection{The Classification of the finite subgroups of
    \texorpdfstring{$\mathrm{SO}(3)$}{} (1.8)}
    \label{sec:Classif-subg-SO3}

@@

    This equation can be simplified by considering the practical numerical values of
    $n$, $r$ and $n_i$, with $n_i \leq n$. By eliminating case by case, he finally arrived
    at five sets of possible values:
    \begin{table}[H]
    \centering
    \label{tab:Finite-rot-g}
    \caption{Finite rotation groups}
    \begin{tabular}{l l l l l l}
        \hline
        $r$ & $(n_1,n_2,n_3)$ & $\# G$ & Schoenflies & Hermann-Mauguin & Note
        \\
        \hline
        \hline
        2 & (n,n,0) & n & $C_n$ & $n$ & Cyclic group \\
        3 & $(2,2,k), k\geq 2$ & $2k$ & $D_k$ & $222$ for $D_2$, $k2$ otherwise & Dihedral group \\
        3 & $(2,3,3)$ & 12  & $T$  & 23& of regular tetrahedron \\
        3 & $(2,3,4)$ & 24 & $O$ & 432 & of regular octahedron \\
        3 & $(2,3,5)$ & 60 & $I$ & not mentioned & of icosahedron, of
        Fullerene\footnotemark\\
        \hline
    \end{tabular}
    \end{table}
    \footnotetext{See section 1.10 about Fullerene and icosahedron for details.}
    For details about derivation please visit pp. 28 to 31 of \cite{Sternberg1994}.
    But we need to exclude some subgroups from the list of crystallographic groups.
    The result is that, on the first row, $n$ is restricted to be $1,2,3,4$ and $6$;
    on the second row, $k$ is restricted to be $2,3,4$ and $6$. And $I$ is excluded
    from the list of crystallographic groups.

@@

    \subsection{Classification of finite subgroups of
    \texorpdfstring{$\mathrm{O}(3)$}{}}
    \label{sec:Classif-subg-O3}

    This section based on the work of previous
    section~\ref{sec:Classif-subg-SO3}. To use the result of those groups of
    matrices with $\det=1$ in that section, the author first discuss the
    example of $T_d$ and $O$.

    $T_d$ is the group of all symmetries(including those that does not preserve
    the chirality, i.e. with $\det=-1$) of tetrahedron. As abstract groups,
    $T_d$ and $O$ are all isomorphic to $S_4$, the symmetric group of order $4$.
    Note that in the context of crystal symmetry, the isomorphic class of group is
    not enough to characterize a behavior. For example, the $C_2$
    containing $\pi$ rotation is not the same as $\{e,i\}$ where $i$ is
    the inversion, but they are isomorphic.\footnote{Perhaps representation
    classes of groups is more precise in describing the behavior.} The link
    between $T_d$ and $O$ can be seen from the following:

    \begin{myquote} \enquote{
    we can consider our tetrahedron as situated inside a cube, with its vertices
    at the diagonally opposite vertices of each face of the cube. For a given
    cube there are two such tetrahedra, as shown in Fig. 1.13
    (Figure~\ref{fig:fig1.13} here).
    One tetrahedron is carried into the other by the inversion, $-I$.
    } \end{myquote}
    \begin{figure}[H]
        \centering
        \includegraphics[width=0.6\linewidth]{pics/{fig1.13}.png}
        \caption{Fig 1.13 in \cite{Sternberg1994}, with modification}
        \label{fig:fig1.13}
    \end{figure}

    Written mathematically, we have
    \begin{equation}
        T_d = S_4 = T\cup T'
    \end{equation}
    where $T$ is mentioned in Table~\ref{tab:Finite-rot-g}. And $T'$ contains
    those matrices with $\det=-1$.
    \begin{equation}
        T_d \overset{\phi}{\cong} O
    \end{equation}
    where $\phi$ is defined as
    \begin{align*}
        A \to A\quad &\text{for }A\in T \\
        B \to -B\quad &\text{for }B\in T'
    \end{align*}

    In general case, for a finite subgroup $G$ of $\mathrm{O}(3)$, we have the
    following. Let $G_+$ denote the subgroup of $G$ with positive determinant.
    Let $G_-$ denote the set of elements of $G$ with negative determinant. We
    have two cases.

    Case 1, $-I\in G_-$, then $G_-=(-I)G_+$, and the structure is simple.

    Case 2, $-I\notin G_-$. Then the author defines $G^\vee:=G_+\cup (-I)G_-$,
    and shows that 
    \begin{align}
        G^\vee &= G_+ \cup aG_+ \\
        G_- &= (-I)aG_+
    \end{align}
    where $a$ is some matrix in $G^\vee\setminus G_+$.

    Then,
    \begin{myquote} \enquote{
    From the $11$ rotation groups listed at the end of the preceding section, we
    obtain $11$ non-rotation groups by including $-I$. In addition, we must check for
    the normat subgroups of index two in each rotation group in order to
    construct non-rotation groups which do not include $-I$.
    } \end{myquote}

    The rest of this section carries out this analysis, culminating in the
    table~ 4 on pp. 40 of \cite{Sternberg1994}. I reproduce the table with some
    addition in the Table~\ref{tab:point-g-32}.

    \begin{table}[H]
        \centering
        \caption{The $32$ point groups}
        \label{tab:point-g-32}
        \includegraphics[width=1\linewidth]{pics/table4-32-point-group.png}
    \end{table}

    The detailed derivation could be found on pp.35 to 39 of
    \cite{Sternberg1994}. Note that $S^6\neq S_6$, and $S^4\neq S_4$. Here is
    some explanations for subscripts: 
    \begin{table}[H]
    \centering
    \begin{tabular}{c l}
        $h$: & there is a horizontal reflection plane. \\
        $v$: & there is a vertical reflection plane. \\
        $d$: & there is diagonal vertical planes. \\
        $m$: & there is a vertical mirror plane (in H-M notation). \\
        $/m$: & there is a horizontal mirror plane (in H-M notation).
    \end{tabular}
    \end{table}

    Note also that the online engine WolframAlpha is a very good source for
    details about certain groups. For example, this
    \href{http://m.wolframalpha.com/input/?i=crystallographic+point+group+C4h}{link}
    is a good description of group $S^4$ (thouhg it is denoted $S_4$ on the
    webpage). Searching "crystallographic point group C4h" will get one a good
    \href{http://m.wolframalpha.com/input/?i=crystallographic+point+group+C4h}{description}
    of $C_{4h}$. One can also found the book reference from that page by
    clicking the "Source information" link.

    Then the author goes on to mention some historical methods in "determination
    of the actual symmetry groups of a given substance", "available to 19th
    century crystallographers". One method is "to attack the crystal with an
    acid, and to study the etch figures induced on the faces".  Another is to
    study the "piezoelectricity", or to test the "rotatory optical activity".

    He also gives a helpful graphical display of $32$ point groups in Fig 1.19
    on \cite{Sternberg1994}(from pp.42 to 44), which is too long to be included
    here.
\bibliography{cite}
\bibliographystyle{alpha}
